---
title: "ce_supplemental_data"
description: reading supplemental prescription data from CMS.gov, plot, publish to Plotly
output: html_document
---


```{r}
library(tidyverse)
library(dplyr)
library(plotly)
library(ggplot2)
library (readr)
library(httr)
library(stringr)
library(RSocrata)
```

#Query the Medicare Prescriber data API
https://github.com/chicago/rsocrata
https://cran.r-project.org/web/packages/RSocrata/RSocrata.pdf

```{r}
#Load Socrata token
Sys.getenv("token")

token <- "wjfjU7dJlpvVZ3nAOq7goSfin"
```

#Drug Summary Nat'l Statistics
2017 Drug summary: https://dev.socrata.com/foundry/data.cms.gov/szqx-nrkt
```{r}
drug_summary<- read.socrata("https://data.cms.gov/resource/szqx-nrkt.csv", app_token = token)

drug_summary_2017 <- drug_summary

nrow(drug_summary_2017)
```

#Drug Summary TN Statistics
```{r}

tn_drug_summary_15 <- read.socrata("https://data.cms.gov/resource/5f7p-yw36.csv?$where=state='Tennessee'", app_token = token) %>%
  mutate(year=2015)
glimpse(tn_drug_summary_15)
nrow(tn_drug_summary_15)

tn_drug_summary_16 <- read.socrata("https://data.cms.gov/resource/hjv3-puc7.csv?$where=state='Tennessee'", app_token = token)%>%
  mutate(year=2016)
glimpse(tn_drug_summary_16)
nrow(tn_drug_summary_16)

tn_drug_summary_17 <- read.socrata("https://data.cms.gov/resource/rrr7-49mp.csv?$where=state='Tennessee'", app_token = token)%>%
  mutate(year=2017)

glimpse(tn_drug_summary_17)
nrow(tn_drug_summary_17)

tn_drug_summary_18 <- read.socrata("https://data.cms.gov/resource/y285-qrb6.csv?$where=state_name='Tennessee'", app_token = token)%>%
  mutate(year=2018)%>%
  rename(state=state_name)

glimpse(tn_drug_summary_18)
nrow(tn_drug_summary_18)

colnames(tn_drug_summary_18) <- colnames(tn_drug_summary_17) #standardize colnames

#Combine dataframes for 2015-2018
tn_drug_summary <- bind_rows(tn_drug_summary_15, tn_drug_summary_16, tn_drug_summary_17, tn_drug_summary_18)
glimpse(tn_drug_summary)
```


#TN Prescriber summary
2017 Prescriber summary : https://dev.socrata.com/foundry/data.cms.gov/psut-35i4
```{r}
prescriber_summary <- read.socrata("https://data.cms.gov/resource/psut-35i4.csv?$where=nppes_provider_state='TN'", app_token = token)

prescriber_summary_2017 <- prescriber_summary
nrow(prescriber_summary_2017)

```

#TN Provider utilization

2017 Provider utilization: https://dev.socrata.com/foundry/data.cms.gov/77gb-8z53
```{r}
#path: https://data.cms.gov/resource/psut-35i4.csv?$where=nppes_provider_state="TN"

provider_utilization <- read.socrata("https://data.cms.gov/resource/77gb-8z53.csv?$where=nppes_provider_state='TN'", app_token = token)

provider_utilization_2017 <- provider_utilization #rename for year
nrow(provider_utilization_2017)

```

```{r}
drug_summary%>%colnames()

colname_cleanup <- function(x) {
  colnames(x) <- x %>%
    colnames() %>%
    str_to_lower() %>% #make colnames lowercase
    str_replace_all("-", "_") %>% #replace hyphens with underscores
    str_replace_all(" ", "_") %>% #replace spaces with underscores
    str_replace_all("\\/", "_")%>% #replace slashes with underscores
    str_replace_all("\\/", "_")
  ;x 
}
```

#Calculate prescription rates by opioid
```{r}
glimpse(tn_drug_summary)

tn_opioid_prescription_rates <- tn_drug_summary%>%
  filter(opioid_drug_flag=="Y")%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries,
         year
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(generic_name, year)%>%
  arrange(generic_name, year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))

view(tn_opioid_prescription_rates)

tn_opioid_prescription_rates%>%
  mutate(generic_name=generic_name%>%
  str_replace("ACETAMINOPHEN WITH CODEINE|ACETAMINOPHEN\\/CAFF\\/DIHYDROCOD|BUTALBIT\\/ACETAMIN\\/CAFF\\/CODEINE|CODEINE SULFATE|CODEINE\\/BUTALBITAL\\/ASA\\/CAFFEIN|DHCOEINE BT\\/ACETAMINOPHN\\/CAFF", "CODEINE")%>%
    str_replace("BUPRENORPHINE|BUPRENORPHINE HCL", "BUPRENORPHINE")%>%
    str_replace("FENTANYL|FENTANYL CITRATE", "FENTANYL")%>%
    str_replace("HYDROCONEONE BITARTRATE|HYDROCODONE\\/ACETAMINOPHEN|HYDROCODONE\\/IBUPROFEN", "HYDROCODONE")%>%
    str_replace("MORPHINE SULFATE|MORPHINE SULFATE\\/NALTREXONE", "MORPHINE")%>%
    str_replace("OXYCODONE HCL|OCYCODONE HCL\\/ACETAMINOPHEN\OXYCODONE HCL\\/ASPIRIN|OXYCODONE MYRISTATE", "OXYCODONE")%>%
    str_replace("PENTAZOCINE HCL\\/NALOXONE HCL","PENTAZOCINE")%>%
    str_replace("TRAMADOL HCL|TRAMADOL HCL\\/ACETAMINOPHEN")%>%
    str_remove(" HCL"))

tn_opioid_prescription_rates%>%
  distinct(generic_name)
```

#Summary metrics for TN Medicare opioid claims 2015-2018
```{r}
tn_drug_summary%>%
  filter(opioid_drug_flag=="Y")%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries,
         year
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(year)%>%
  arrange(year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))
```
#Aggregate generic opioid drug names for plotting
```{r}

tn_opioid_rate_summary <- tn_drug_summary%>%
  filter(opioid_drug_flag=="Y")%>%
  mutate(generic_name=generic_name%>%
           str_replace("DHCODEINE BT\\/ACETAMINOPHN\\/CAFF", "CODEINE")%>%
           str_replace("ACETAMINOPHEN WITH CODEINE","CODEINE")%>%
           str_replace("ACETAMINOPHEN\\/CAFF\\/DIHYDROCOD","CODEINE")%>%
           str_replace("BUTALBIT\\/ACETAMIN\\/CAFF\\/CODEINE","CODEINE")%>%
           str_replace("BUTORPHANOL TARTRATE", "BUTORPHANOL")%>%
           str_replace("CODEINE SULFATE","CODEINE")%>%
           str_replace("CODEINE\\/BUTALBITAL\\/ASA\\/CAFFEIN","CODEINE")%>%
           str_replace("BUPRENORPHINE HCL", "BUPRENORPHINE")%>%
           str_replace("FENTANYL CITRATE", "FENTANYL")%>%
           str_replace("HYDROCODONE BITARTRATE", "HYDROCODONE")%>%
           str_replace("HYDROCODONE\\/ACETAMINOPHEN|HYDROCODONE\\/IBUPROFEN", "HYDROCODONE") %>%
           str_replace("MORPHINE SULFATE\\/NALTREXONE", "MORPHINE")%>%
           str_replace("MORPHINE SULFATE", "MORPHINE")%>%
           str_replace("OXYCODONE HCL\\/ASPIRIN", "OXYCODONE")%>%
           str_replace("OXYCODONE HCL", "OXYCODONE")%>%
           str_replace("OXYCODONE MYRISTATE", "OXYCODONE")%>%
           str_replace("OXYCODONE\\/ACETAMINOPHEN", "OXYCODONE")%>%
           str_replace("PENTAZOCINE HCL\\/NALOXONE HCL","PENTAZOCINE")%>%
           str_replace("TRAMADOL HCL\\/ACETAMINOPHEN", "TRAMADOL")%>%
           str_remove(" HCL")
         )%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries,
         year
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(generic_name, year)%>%
  arrange(generic_name, year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))

view(tn_opioid_rate_summary)

```

#Plot changes in TN Medicaid opioid prescription claim volume 2015-2018
Stacked bar chart
```{r}
p <- tn_opioid_rate_summary%>% 
          filter(generic_name!="LEVORPHANOL TARTRATE")%>% #Exclude very low volume opioids
          filter(generic_name!="OPIUM/BELLADONNA ALKALOIDS")%>% #This drug used for treating spasms in urinary tract
          rename(`Opioid` = generic_name,
                     `# Medicare Claims` = num_claims,
                     Year = year) %>%
  ggplot(aes(fill=`Opioid`, 
           y=`# Medicare Claims`, 
           x=Year)
       ) +
    geom_bar(position="stack", stat="identity")+ 
  scale_y_continuous(labels = scales::comma) 
  
fig<-ggplotly(p)

fig

#publish to Plotly
#api_create(fig, filename = "opioid_rate_decrease")
```
Time series
```{r}

fig <- tn_opioid_rate_summary%>% 
  filter(generic_name!="LEVORPHANOL TARTRATE")%>%
  filter(generic_name!="OPIUM/BELLADONNA ALKALOIDS")%>%
  ggplot( aes(x=year, 
              y=num_claims, 
              color=generic_name)) +
    geom_line()

ggplotly(fig)


```
#Examine specific forumlations of Naloxone for frequency of emergency injections/nasal mist claims
```{r}

naloxone <- tn_drug_summary%>%
  filter(str_detect(generic_name, 'NALOXONE'))%>%
  select(drug_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries,
         year
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(drug_name, year)%>%
  arrange(drug_name, year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))

view(naloxone)
```

#Separate claim data for medically assisted opioid use disorder treatment
```{r}

oud_treatments <- tn_drug_summary%>%
  filter(str_detect(generic_name, 'METHADONE|NALTREXONE|NALOXONE|BUPRENORPHINE'))%>%
  mutate(generic_name=generic_name%>%
           str_replace("BUPRENORPHINE HCL/NALOXONE HCL","BUPRENORPHINE")%>%
           str_replace("BUPRENORPHINE HCL", "BUPRENORPHINE")%>%
           str_replace("METHYLNALTREXONE BROMIDE","NALTREXONE")%>%
           str_replace("MORPHINE SULFATE/NALTREXONE", "NALTREXONE")%>%
           str_replace("NALTREXONE MICROSPHERES","NALTREXONE")%>%
           str_replace("PENTAZOCINE HCL/NALOXONE HCL","NALOXONE")%>%
           str_remove(" HCL")
         )%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries,
         year
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(generic_name, year)%>%
  arrange(generic_name, year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))

view(oud_treatments)

oud_treatments%>%distinct(generic_name)

```

OUD medicine claims summarised by year
```{r}
oud_treatments%>%group_by(year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))
```

#PLOT OUD TREATMENTS

```{r}
#Stacked bar
fig <- oud_treatments%>% #tn_opioid_prescription_rates%>%
  rename(`Generic Name` = generic_name,
                     `# Medicare Claims` = num_claims,
                     Year = year) %>%
  ggplot( aes(x=Year, 
              y=`# Medicare Claims`, 
              #group=fipscounty, 
              color=`Generic Name`)) +
    geom_line(size=1.25) +
    geom_point(shape=21, size=2)
  #geom_bar(position="stack", stat="identity")

ggplotly(fig)

#Publish to plotly
#api_create(fig, filename = "oud_treatment_claims")
```

```{r}
oud_treatments_grouped <- tn_drug_summary%>%
  filter(str_detect(generic_name, 'METHADONE|NALTREXONE|NALOXONE|BUPRENORPHINE'))%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries,
         year
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(generic_name, year)%>%
  arrange(generic_name, year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))

oud_treatments_grouped
```

```{r}
oud_treatments_sum <- tn_drug_summary%>%
  filter(str_detect(generic_name, 'METHADONE|NALTREXONE|NALOXONE|BUPRENORPHINE'))%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries,
         year
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(year)%>%
  arrange(year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))

view(oud_treatments_sum)
```


#Summary changes
```{r}
tn_opioid_prescription_rates_sum <- tn_drug_summary%>%
  filter(opioid_drug_flag=="Y")%>%
#  select(generic_name, 
#         number_of_prescribers, 
#         number_of_medicare_part_d_claims,
#         number_of_medicare_beneficiaries,
#         year
#         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries,
         num_claims_ge65 = number_of_medicare_part_d_claims_for_beneficiaries_65,
         num_beneficiaries_ge65 = number_of_medicare_beneficiaries_65)%>%
  group_by(year)%>%
  arrange(year)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE),
            num_claims_ge65=sum(num_claims_ge65,na.rm = TRUE),
            num_beneficiaries_ge65=sum(num_beneficiaries_ge65,na.rm = TRUE)
            )%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2),
         claim_per_benficiary_ge65 = round ((num_claims_ge65/num_beneficiaries_ge65),2))

tn_opioid_prescription_rates_sum
colnames(tn_drug_summary)
```


```{r}
glimpse(drug_summary_2017)

opioid_prescription_rates_2017 <- drug_summary_2017%>%
  filter(opioid_drug_flag=="Y")%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  group_by(generic_name)%>%
  arrange(generic_name)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))


drug_summary_2017%>%
  filter(opioid_drug_flag=="Y")%>%
  select(generic_name, 
         number_of_prescribers, 
         number_of_medicare_part_d_claims,
         number_of_medicare_beneficiaries
         )%>%
  rename(num_prescribers = number_of_prescribers,
         num_claims = number_of_medicare_part_d_claims,
         num_beneficiaries = number_of_medicare_beneficiaries)%>%
  summarise(num_prescribers=sum(num_prescribers, na.rm = TRUE),
            num_beneficiaries = sum(num_beneficiaries,na.rm = TRUE),
            num_claims=sum(num_claims,na.rm = TRUE))%>%
  mutate(claim_per_beneficiary = round((num_claims/num_beneficiaries),2),
         claim_per_prescriber = round((num_claims/num_prescribers),2))


view(opioid_prescription_rates_2017)
```

```{r}
oud_treatments <- tn_drug_summary%>%
  filter(str_detect(generic_name, 'METHADONE|NALTREXONE|NALOXONE|BUPRENORPHINE'))

oud_treatments
```

```{r}
prescriber_summary_2017%>%colnames()

prescriber_risk_2017 <- prescriber_summary_2017%>%select(npi,
         nppes_provider_last_org_name,
         nppes_provider_first_name,
         specialty_description,
         bene_count,
         total_claim_count,
         opioid_prescriber_rate,
         opioid_claim_count,
         opioid_bene_count)%>%
  mutate(opioids_per_beneficiary = round((opioid_claim_count/opioid_bene_count),2))

view(prescriber_risk_2017)
```

